\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-04-18 doc/paper.Rmd
define q_i

The Chumbley score algorithm takes input in form of two digitized toolmarks. The toolmark is in form of $z(t)$ which is a spatial process for location indexed by $t$. $t$ here denotes equally spaced pixel locations for the striation marks under consideration, $t = 1, ..., T$. Let further $z(s,t)$ denote the vector of markings between locations $s$ and $t$.

@@

After removing sub-class structure, 
the Chumbley scores is calculated in two steps: an optimization step and a validation step.
In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
\[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
\]
This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings. \hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$. Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$. }

@@

\hh{XXX summarise in formula}

@@

We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans provided by NIST (a total of 85,491 comparisons). The settings for optimizing and validating window sizes, $w_o$ and $w_v$, ranged from $w_o \in [50, 280]$ and $w_v \in \{30, 50\}$, see also figure \ref{fig:type2}.

@@

\hh{number of failed tests for default parameter settings -- XXX what are the defaults? wo:10 percent, coarseness 0.25, wv XXX?}

@@

p2 <- profc %>% 
  ggplot(aes(x = locations/(length-wo))) +
  geom_histogram(binwidth = 0.025) +
  facet_grid(.~coarseness, labeller="label_both") + 
  theme_bw() +
  xlab(expression("Relative optimal location "~t^*)) +
  ylab("Number of profiles")

>>>>

The Chumbley score algorithm takes input in form of two digitized toolmarks. The toolmark is in form of $z(t)$ which is a spatial process for location indexed by $t$. $t$ here denotes equally spaced pixel locations for the striation marks under consideration, $t = 1, ..., T$. Let further $z^s(t)$ denote a vector of markings of length $s$ starting in location  $t$.

@@

After removing sub-class structure, 
the Chumbley scores is calculated in two steps: an optimization step and a validation step.
In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
\[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
\]
This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings. \hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$. Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$. }

@@

More precisely, \citet{hadler} define same-shift pairs of length $w_v$ to start in $(q^{(1)}_i, q^{(2)}_i)$, where $q^{(k)}_i$ has the form
\begin{eqnarray*}
q^{(k)}_i &=& t_k^* + i w_v \text{ for } i = -1, -2, ...\\
q^{(k)}_i &=& t_k^* + w_ o + i w_v \text{ for } i = 0, 1, 2, ...
\end{eqnarray*}
for all integer values of $i$ for which both $0 <  q^{(1)}_i \le T_1 - w_v$ and $0 <  q^{(2)}_i \le T_2 - w_v$.

Similarly, different-shift pairs are based on signature segments of length $w_v$ starting in $(q^{(1)}_i, q^{(2)}_{-i})$, for all integer values of $i$ for which both $0 <  q^{(1)}_i \le T_1 - w_v$ and $0 <  q^{(2)}_{-i} \le T_2 - w_v$.

@@

We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans provided by NIST for a total of 85,491 land-to-land comparisons.

<!--The settings for optimizing and validating window sizes, $w_o$ and $w_v$, ranged from $w_o \in [50, 280]$ and $w_v \in \{30, 50\}$, see also figure \ref{fig:type2}.-->

@@

Initially, the default settings suggested in \citet{hadler} were used: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$. \autoref{fig:fails} shows the percentage of failed tests among the 86k land-to-land comparisons of the NIST data for different values of the validation window size $w_v$. For same-source lands up to 12.5 percent of the tests fail.

```{r fails, echo=FALSE, fig.height = 3, fig.width = 5, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$. With an increase in the size of the validation window  a higher percentage of tests fails. The percentage of failures is higher for same-source lands than for different source lands."}
fails <- allp %>% filter(coarse == 0.25) %>% 
  group_by(wo, wv) %>% summarize(
  failed = sum(is.na(p_value)),
  n=n()
)

allp %>% filter(coarse == 0.25) %>% 
  group_by(wo, wv, match) %>% summarize(
  failed = sum(is.na(p_value)),
  n=n()
) %>% 
  ggplot(aes(x = wv, y = failed/n*100, colour=factor(match), shape=factor(match))) +
  geom_point(size=3) +
  xlab("Size of validation window in pixels") +
  ylab("Percent of failed tests") +
  theme_bw() +
  scale_colour_brewer("Same-source", palette="Set1") +
  scale_shape_discrete("Same-source") +
  ylim(c(0,13))
```

@@

p2 <- profc %>% 
  ggplot(aes(x = locations/(length-wo))) +
  geom_histogram(binwidth = 0.025) +
  facet_grid(.~coarseness, labeller="label_both") + 
  theme_bw() +
  xlab(expression("Relative optimal location "~t^"*")) +
  ylab("Number of profiles")

### 2018-04-18 doc/paper.tex
define q_i

The Chumbley score algorithm takes input in form of two digitized
toolmarks. The toolmark is in form of \(z(t)\) which is a spatial
process for location indexed by \(t\). \(t\) here denotes equally spaced
pixel locations for the striation marks under consideration,
\(t = 1, ..., T\). Let further \(z(s,t)\) denote the vector of markings
between locations \(s\) and \(t\).

@@

After removing sub-class structure, the Chumbley scores is calculated in
two steps: an optimization step and a validation step. In the
optimization step, the two markings are aligned horizontally such that
within a pre-defined window of length \(w_o\) the correlation between
\(x(t_1)\) and \(y(t_2)\) is maximized: \[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
\] This results in an optimal vertical (in-phase) shift of
\(t_1^o - t_2^o\) for aligning the two markings.
\hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$. Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$. }

@@

\hh{XXX summarise in formula}

@@

We used the adjusted Chumbley method as proposed in \citet{hadler} and
implemented in the R package \texttt{toolmaRk} \citep{toolmark} on all
pairwise land-to-land comparisons of the Hamby scans provided by NIST (a
total of 85,491 comparisons). The settings for optimizing and validating
window sizes, \(w_o\) and \(w_v\), ranged from \(w_o \in [50, 280]\) and
\(w_v \in \{30, 50\}\), see also figure \ref{fig:type2}.

\section{Results}\label{results}

@@

\subsection{Results}\label{results-1}

>>>>

The Chumbley score algorithm takes input in form of two digitized
toolmarks. The toolmark is in form of \(z(t)\) which is a spatial
process for location indexed by \(t\). \(t\) here denotes equally spaced
pixel locations for the striation marks under consideration,
\(t = 1, ..., T\). Let further \(z^s(t)\) denote a vector of markings of
length \(s\) starting in location \(t\).

@@

After removing sub-class structure, the Chumbley scores is calculated in
two steps: an optimization step and a validation step. In the
optimization step, the two markings are aligned horizontally such that
within a pre-defined window of length \(w_o\) the correlation between
\(x(t_1)\) and \(y(t_2)\) is maximized: \[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
\] This results in an optimal vertical (in-phase) shift of
\(t_1^o - t_2^o\) for aligning the two markings.
\hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$. Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$. }

@@

More precisely, \citet{hadler} define same-shift pairs of length \(w_v\)
to start in \((q^{(1)}_i, q^{(2)}_i)\), where \(q^{(k)}_i\) has the form

\begin{eqnarray*}
q^{(k)}_i &=& t_k^* + i w_v \text{ for } i = -1, -2, ...\\
q^{(k)}_i &=& t_k^* + w_ o + i w_v \text{ for } i = 0, 1, 2, ...
\end{eqnarray*}

for all integer values of \(i\) for which both
\(0 < q^{(1)}_i \le T_1 - w_v\) and \(0 < q^{(2)}_i \le T_2 - w_v\).

Similarly, different-shift pairs are based on signature segments of
length \(w_v\) starting in \((q^{(1)}_i, q^{(2)}_{-i})\), for all
integer values of \(i\) for which both \(0 < q^{(1)}_i \le T_1 - w_v\)
and \(0 < q^{(2)}_{-i} \le T_2 - w_v\).

@@

We used the adjusted Chumbley method as proposed in \citet{hadler} and
implemented in the R package \texttt{toolmaRk} \citep{toolmark} on all
pairwise land-to-land comparisons of the Hamby scans provided by NIST
for a total of 85,491 land-to-land comparisons.

\subsection{Failed Tests}\label{failed-tests}

Initially, the default settings suggested in \citet{hadler} were used:
\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
shows the percentage of failed tests among the 86k land-to-land
comparisons of the NIST data for different values of the validation
window size \(w_v\). For same-source lands up to 12.5 percent of the
tests fail.

{\centering \includegraphics[width=0.7\textwidth]{figures/fails-1}

\caption{Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$. With an increase in the size of the validation window  a higher percentage of tests fails. The percentage of failures is higher for same-source lands than for different source lands.}\label{fig:fails}
\end{figure}

\hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
\hh{figure: number of diff shift pairs and location}

\section{A modified test}\label{a-modified-test}

\hh{fix the algorithm - all of the other properties in Hadler and morris still hold.}

\hh{number of failed tests down. }

\section{Comparison of Results}\label{comparison-of-results}

@@

\subsection{Results}\label{results}
